---
title: "R Group Project Daniel"
output: pdf_document
date: "2023-12-08"
---

```{r}
require(tidyverse)
require(lme4)
```

# Raw Data
```{r}
Raw_data <- read.csv2("dataset_Rubalcaba_Polo.csv")
```

# Data Structure
```{r}
str(Raw_data) #The structure of the data
colnames(Raw_data) #It's column names
```

# Organizing Data
```{r}
columns_used_data <- Raw_data %>% 
select(Eggs, Hatchlings, Fledglings, Males, Females, Year) # Columns we will use
```

```{r}
data_with_added_columns <- columns_used_data %>%
  mutate(
    deaths = Eggs - Fledglings,
    survival_rate = Fledglings/Eggs,
    percent_female = Females/(Males + Females)
  ) #Adding all the columns we will use for analysis
```

# Data Explaination
```{r}
str(data_with_added_columns)
```
Egg: The amount of starling eggs in each observation
Hatchlings: The amount of hatchlings in each observation
Fledglings: The amount of fledglings in each observation
Males: Amount of males per observation
Females: Amount of females per observation
Year: The year the observation takes place
deaths: Taken as Eggs - Fledglings, is the amount of eggs that didn't survive to fledglings
survival rate: Taken as Fledgling/Egg, the percentage that survived
percent_females: Taken as Females/(Females + Males), the percentage of offsping that is female

# Quick Analysis

## Female Percentage on Survival Rates and Deaths

```{r}
#Using density plot to visually check for normality
ggplot(data_with_added_columns, aes(x = survival_rate)) + 
  geom_density()

ggplot(data_with_added_columns, aes(x = deaths)) + 
  geom_density()
```

Distribution seems non normal from this density plot.

```{r}
data_with_added_columns %>% 
  ggplot(aes(x = percent_female, y = survival_rate)) +
  geom_point() +
  stat_smooth(method = "lm", se = F, colour = "gray")

#Visually check for linearity
```

Points are scattered far from best fit, so it doesn't seem linear.

```{r}
data_with_added_columns %>% 
  ggplot(aes(x = percent_female, y = mean(survival_rate))) +
  geom_point() +
  stat_smooth(method = "lm", se = F, colour = "gray")
#Visually check for linearity using the mean
```

A very flat line, doesn't seem linear.

```{r}
data_with_female_percentage_group <- data_with_added_columns %>% 
 mutate(Female_percentage_group = cut(percent_female,
                                       breaks = seq(0, 1, by = 0.25),  # Bins at 0%, 25%, 50%, 75%, 100%
                                       include.lowest = TRUE,
                                       labels = c("0-25%", "26-50%", "51-75%", "76-100%"))) #Grouping into brackets to see the trend, as to make sure some mean values aren't from one observation

data_with_mean_survival_female_percentage <- data_with_female_percentage_group %>% 
  group_by(Female_percentage_group) %>%
  summarize(Mean_survival_rate = mean(survival_rate, na.rm = TRUE))

survival_rate_plot <- data_with_mean_survival_female_percentage %>% 
  ggplot(aes(x = Female_percentage_group, y = Mean_survival_rate)) +
  geom_point() +
  labs(x = "Female Percentage Group", y = "Mean Survial Rate", title = "Female Percentage by Mean Survial Rate")
#Plot of the trends, using female percentage groups

ggsave("Female Plot.png", plot = survival_rate_plot, width = 10, height = 8, dpi = 300)

survival_rate_plot
```

Seems to be a trend of positive slope till around 75%, then negative.

```{r}
data_with_mean_death <- data_with_female_percentage_group %>% 
  group_by(Female_percentage_group) %>%
  summarize(Mean_deaths = mean(deaths, na.rm = TRUE))

data_with_mean_death %>% 
  ggplot(aes(x = Female_percentage_group, y = Mean_deaths)) +
  geom_point() +
  labs(x = "Female Percentage Group", y = "Mean Deaths", title = "Female Percentage by Mean Deaths")
#Same thing, but with deaths instead
```

It seems that female percentage stops having a positive effect after the 51-75% group on survival rate and death.

# Model of #Female to #Fledgling

Since the data doesn't seem to be normal, or linear, GLM will be used instead.

```{r}
model_poisson <- glm(Fledglings ~ Females, family = poisson, data = data_with_added_columns) #The linear Poisson model

summary(model_poisson)
```

```{r}
model_poisson_quadratic <- glm(Fledglings ~ Females + I(Females^2), family = poisson, data = data_with_added_columns) #The quadratic Poisson model

summary(model_poisson_quadratic)
```

The AIC score differs by <2, so the models are indistinguishable in this regard.

From the linear model, it seems there is a significance of Amount of Females in a nest and the amount of eggs that survive into Fledglings, with a p value of 9.91e-06.

## Graphing the model
```{r}
Range_of_Females <- seq(from = min(data_with_added_columns$Females), to = 10, by = 1) #Upping the range to 10 to see how the model continues

Prediction <- data.frame(Females = Range_of_Females) #Making it into a data frame

Prediction$Fledglings_Linear <- predict(model_poisson, newdata = Prediction, type = "response")
Prediction$Fledglings_Quadratic <- predict(model_poisson_quadratic, newdata = Prediction, type = "response")
```

```{r}
Model_comparison_plot <- ggplot(data_with_added_columns, aes(x = Females, y = Fledglings)) +
  geom_point(colour = "grey") +
  geom_line(data = Prediction, aes(x = Females, y = Fledglings_Linear), colour = "blue") +
  geom_line(data = Prediction, aes(x= Females, y = Fledglings_Quadratic), colour = "red") +
  labs(title = "Model Comparison: Linear vs Quadratic") #Making the Graph

ggsave("compasison plot.png", plot = Model_comparison_plot, width = 10, height = 8, dpi = 300) #Saving as a png

Model_comparison_plot
```
